---
title: "Tennis_Rankings_Table"
author: "Eli"
date: "7/5/2022"
output: html_document
---

Load in libraries
```{r}
library(tidyverse)
library(lubridate)
library(gt)
library(gtExtras)
```
#Load in data from Jeff Sackmann github for wta AND atp
Rankings
```{r}
wta_rankings <- read_csv(url('https://raw.githubusercontent.com/JeffSackmann/tennis_wta/master/wta_rankings_current.csv'))
wta_rankings$ranking_date <- ymd(wta_rankings$ranking_date)

#Players list
wta_players <- read_csv(url('https://raw.githubusercontent.com/JeffSackmann/tennis_wta/master/wta_players.csv'))
wta_players$Name <- paste(wta_players$name_first, wta_players$name_last, sep = " ")

#merge player names onto rankings
wta_rankings <- wta_rankings %>% 
  mutate(Name = wta_players$Name[match(player, wta_players$player_id)])

#2022 stats
wta_matches_2022 <- data.frame(read.csv(url('https://raw.githubusercontent.com/JeffSackmann/tennis_wta/master/wta_matches_2022.csv')))

#sort matches oldest to most recent
wta_matches_2022$tourney_date <- ymd(wta_matches_2022$tourney_date)
wta_matches_2022 <- wta_matches_2022 %>% 
  arrange(tourney_date,match_num) %>% 
  drop_na(c(winner_rank,loser_rank))
```

retrieve current top 10, top 15, etc
```{r}
wta_top_ten <- wta_rankings %>% 
  filter(ranking_date >= max(ranking_date) & rank <= 10)
```

Retrieve player's best win and worst loss
Let's write a function! Input player name and spit out their worst loss and best win
```{r}
best_and_worst_result <- function(Player_Name){
 
  wins <- wta_matches_2022 %>% 
    filter(winner_name == Player_Name)
  
  best_win_rank <- min(wins$loser_rank)
  best_win_name <- wins$loser_name[wins$loser_rank == best_win_rank]
  best_win <- paste(best_win_rank,best_win_name)
  
  losses <- wta_matches_2022 %>% 
    filter(loser_name == Player_Name)
  
  worst_loss_rank <- max(losses$winner_rank)
  worst_loss_name <- losses$winner_name[losses$winner_rank == worst_loss_rank]
  worst_loss <- paste(worst_loss_rank,worst_loss_name)
  
  return(c(best_win,worst_loss))
}
```

Test out function
```{r}
best_and_worst_result('Iga Swiatek')
```

```{r}
#Build quick loop to iterate through the top 10 (or x) players and attach results to their ranking page
for (i in 1:nrow(wta_top_ten)) {
  results <- best_and_worst_result(wta_top_ten$Name[i])
  wta_top_ten$Best_Win[i] <- results[1]
  wta_top_ten$Worst_Loss[i] <- results[2]
}
```

Start with basic gt table, before I try to add fancy stuff
```{r}
wta_top_ten <- wta_top_ten %>% 
  select('rank', 'Name','Best_Win','Worst_Loss')
wta_top_ten <- wta_top_ten %>% 
  mutate(Country = wta_players$ioc[match(Name, wta_players$Name)]) %>% 
  relocate(Country, .after = Name)
```
#Build table woo woo
```{r}
top_ten_table <- wta_top_ten %>% 
  gt()
top_ten_table
#first version of table is ok... I want to get country code in between Name and Best Win to not have all the names right by each other and confuse vieewer... i'll add country code above
```

Let's work on sparklines
```{r}
#First need to groupby player in the rankings and then capture rank
ranklines <- wta_rankings %>%
  mutate(rank = -1*rank) %>% 
  group_by(Name) %>% 
  summarise(Rank_Change = list(rank))
```

Add listed column of ranks to top ten
```{r}
wta_top_ten <- wta_top_ten %>% 
  mutate(Rank_Change = ranklines$Rank_Change[match(Name, ranklines$Name)])
```

Now try and add it to table
```{r}
top_ten_table <- wta_top_ten %>% 
  gt() %>% 
  gt_sparkline(Rank_Change, label = FALSE)

top_ten_table
```

Time to add win/loss ticks
```{r}
#Build a function to create list of their past 10 match results... assign 0 for loss, 1 for win 
#Something like tail(10)
#If winner_name == player_name, assign 1, if not, passing 0 to list
#Build a for loop to loop through each players last 10 matches 

player_form <- function(player_name){
  form_list <- c()
  #get their last 10 matches
  player_matches <- wta_matches_2022 %>% 
    filter(winner_name == player_name | loser_name == player_name) %>% 
    tail(10)
  
  for (i in 1:nrow(player_matches)) {
    if(player_matches$winner_name[i] == player_name){
      form_list[i] <- 1
    }else{
      form_list[i] <- 0
    }
  }
  return(form_list)
}

player_form('Iga Swiatek')
```

Loop through top ten, applying new function to get list of 1's and 0's for each player
```{r}
for (i in 1:nrow(wta_top_ten)) {
  wta_top_ten$Form[i] <- list(player_form(wta_top_ten$Name[i]))
}
```

Now try to add form to table
```{r}
top_ten_table <- wta_top_ten %>% 
  gt() %>% 
  tab_header(title = "WTA Top Ten 2022 Form and Best/Worst Results",
              subtitle = "Swiatek and Jabeur show their run of form") %>% 
  gt_sparkline(Rank_Change, label = FALSE) %>%  
  gt_plt_winloss(Form) %>% 
  cols_label(
    rank = "Rank",
    Best_Win = "Best Win",
    Worst_Loss = "Worst Loss",
    Rank_Change = 'Change in Rank (2022)',
    Form = 'Last 10 Matches') %>% 
  tab_source_note(source_note = "Data: Jeff Sackmann Repo") %>% 
  gt_theme_538()
top_ten_table
```

Save table
```{r}
gtsave_extra(top_ten_table, 'wta_table.png',  path = getwd())

```

Now copy above steps for ATP!
#Load in data from Jeff Sackmann github for wta AND atp
Rankings
```{r}
atp_rankings <- read_csv(url('https://raw.githubusercontent.com/JeffSackmann/tennis_atp/master/atp_rankings_current.csv'))
atp_rankings$ranking_date <- ymd(atp_rankings$ranking_date)

#Players list
atp_players <- read_csv(url('https://raw.githubusercontent.com/JeffSackmann/tennis_atp/master/atp_players.csv'))
atp_players$Name <- paste(atp_players$name_first, atp_players$name_last, sep = " ")

#merge player names onto rankings
atp_rankings <- atp_rankings %>% 
  mutate(Name = atp_players$Name[match(player, atp_players$player_id)])

#2022 stats
atp_matches_2022 <- data.frame(read.csv(url('https://raw.githubusercontent.com/JeffSackmann/tennis_atp/master/atp_matches_2022.csv')))

#sort matches oldest to most recent
atp_matches_2022$tourney_date <- ymd(atp_matches_2022$tourney_date)
atp_matches_2022 <- atp_matches_2022 %>% 
  arrange(tourney_date,match_num) %>% 
  drop_na(c(winner_rank,loser_rank))
```

retrieve current top 10, top 15, etc
```{r}
atp_top_ten <- atp_rankings %>% 
  filter(ranking_date >= max(ranking_date) & rank <= 10)
```

Retrieve player's best win and worst loss
Let's write a function! Input player name and spit out their worst loss and best win
```{r}
atp_best_and_worst_result <- function(Player_Name){
 
  wins <- atp_matches_2022 %>% 
    filter(winner_name == Player_Name)
  
  best_win_rank <- min(wins$loser_rank)[1]
  best_win_name <- wins$loser_name[wins$loser_rank == best_win_rank][1]
  best_win <- paste(best_win_rank,best_win_name)
  
  losses <- atp_matches_2022 %>% 
    filter(loser_name == Player_Name)
  
  worst_loss_rank <- max(losses$winner_rank)[1]
  worst_loss_name <- losses$winner_name[losses$winner_rank == worst_loss_rank][1]
  worst_loss <- paste(worst_loss_rank,worst_loss_name)
  
  return(c(best_win,worst_loss))
}
```

Test out function
```{r}
atp_best_and_worst_result('Carlos Alcaraz')
```

```{r}
#Build quick loop to iterate through the top 10 (or x) players and attach results to their ranking page
for (i in 1:nrow(atp_top_ten)) {
  results <- atp_best_and_worst_result(atp_top_ten$Name[i])
  atp_top_ten$Best_Win[i] <- results[1]
  atp_top_ten$Worst_Loss[i] <- results[2]
}
```

Start with basic gt table, before I try to add fancy stuff
```{r}
atp_top_ten <- atp_top_ten %>% 
  select('rank', 'Name','Best_Win','Worst_Loss')
atp_top_ten <- atp_top_ten %>% 
  mutate(Country = atp_players$ioc[match(Name, atp_players$Name)]) %>% 
  relocate(Country, .after = Name)
```
#Build table woo woo
```{r}
atp_top_ten_table <- atp_top_ten %>% 
  gt()
atp_top_ten_table
#first version of table is ok... I want to get country code in between Name and Best Win to not have all the names right by each other and confuse vieewer... i'll add country code above
```

Let's work on sparklines
```{r}
#First need to groupby player in the rankings and then capture rank
ranklines <- atp_rankings %>%
  mutate(rank = -1*rank) %>% 
  group_by(Name) %>% 
  summarise(Rank_Change = list(rank))
```

Add listed column of ranks to top ten
```{r}
atp_top_ten <- atp_top_ten %>% 
  mutate(Rank_Change = ranklines$Rank_Change[match(Name, ranklines$Name)])
```

Now try and add it to table
```{r}
atp_top_ten_table <- atp_top_ten %>% 
  gt() %>% 
  gt_sparkline(Rank_Change, label = FALSE)

atp_top_ten_table
```

Time to add win/loss ticks
```{r}
#Build a function to create list of their past 10 match results... assign 0 for loss, 1 for win 
#Something like tail(10)
#If winner_name == player_name, assign 1, if not, passing 0 to list
#Build a for loop to loop through each players last 10 matches 

atp_player_form <- function(player_name){
  form_list <- c()
  #get their last 10 matches
  player_matches <- atp_matches_2022 %>% 
    filter(winner_name == player_name | loser_name == player_name) %>% 
    tail(10)
  
  for (i in 1:nrow(player_matches)) {
    if(player_matches$winner_name[i] == player_name){
      form_list[i] <- 1
    }else{
      form_list[i] <- 0
    }
  }
  return(form_list)
}

atp_player_form('Carlos Alcaraz')
```

Loop through top ten, applying new function to get list of 1's and 0's for each player
```{r}
for (i in 1:nrow(wta_top_ten)) {
  atp_top_ten$Form[i] <- list(atp_player_form(atp_top_ten$Name[i]))
}
```

Now try to add form to table
```{r}
atp_top_ten_table <- atp_top_ten %>% 
  gt() %>% 
  tab_header(title = "ATP Top Ten 2022 Form and Best/Worst Results",
              subtitle = "All players have at least a top ten win") %>% 
  gt_sparkline(Rank_Change, label = FALSE) %>%  
  gt_plt_winloss(Form) %>% 
  cols_label(
    rank = "Rank",
    Best_Win = "Best Win",
    Worst_Loss = "Worst Loss",
    Rank_Change = 'Change in Rank (2022)',
    Form = 'Last 10 Matches') %>% 
  tab_source_note(source_note = "Data: Jeff Sackmann Repo") %>% 
  gt_theme_538()
atp_top_ten_table
```

Save table
```{r}
gtsave_extra(atp_top_ten_table, 'atp_table.png',  path = getwd())

```
